---
title: "Build SingleCellExperiment object S000379"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6, 
                      fig.path='./figures/build_SCE/',
                      warning=FALSE, message=FALSE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
suppressPackageStartupMessages({
  library(DropletUtils)
  library(tidyverse)
  library(patchwork)
  library(scran)
  library(scater)
  })

source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# Aim

The poor quality of the fixed nuclei from the Ovarian carcinosaroma PDX experiment prompted an attempt to fix intact fresh cells instead.

2x PDX tumours split into pieces:

Take 25mg from each tumour, mince into pieces and fix for "fix and chop" protocol
To be dissociated with Scott lab enzymes after 4C storage
Take the remainder and dissociate fresh tissue with Sccott lab enzymes 
Collagenase, dispase, DNase
Perform RBC lysis
Fix 2 million cells
The tumours fixed were:

* PDX1264
* PDX1233

The time between fixing the tissue and fixing the dissociated cells was approximately 2 hours.
The sample didn't dissociate fully after 20 min at 37C.
He put the sample for 20 more min at 37C. c.f the fresh tissue takes ~45min.
After 40min the fixed tissue still not dissociating. Anthony filtered with 100uM filter
I resuspended in quenching buffer and added 10% v/v enhancer for 4C storage

## Sample omission

The sample hybridisated with probeset BC2 was low in cell number and very clumpy. Because the last time I ran PDX tumour fixed nuclei the 10x Genomics channel blocked - I omitted this sample.

Therefore only PDX1233 has a matched fix/chop vs fixed dissociated cells

## Capture issue

Daniel Brown
-
Gery Ma of 10x Genomics came onsite to supervise the capture. I counted the pooled cell suspension to ~4,000 cells/uL in 500uL

Targeted 28,000 cells to recover because the countess underestimated the count. After capture was short of GEM volume by ~5uL. There was excess gel bead volume ~40uL

Gerry recommended to capture again with a reimbursed reaction. However had the same issue. Incubated both captures.

More likely in my opinion to be the slight clumping which Gerry said was OK.

## Low librrary concentrration

* Capture 1 = 1.21ng/uL
  + 6.84nM
* Capture 2 = 1.5ng/uL
  + 8.48nM

# Load Data

Read in the filtered sample specific matrices then combine.

```{r capture1}
cap1_samples <- c("PDX1233-cells", "PDX1233-tissue", "PDX1264-cells")
cap1_head <- "/stornext/Projects/score/GenomicsRnD/DB/S000379/R010_PDX_FLEX-1/outs/per_sample_outs/"
tail_path <- "/count/sample_filtered_feature_bc_matrix"

cap1 <- list()
for (i in cap1_samples) {
  path <- paste(cap1_head, i, tail_path, sep = "")
  cap1[[i]] <- read10xCounts(path, col.names = T)
}

cap1[["PDX1233-cells"]]$Sample <- "PDX1233-cells"
cap1[["PDX1233-tissue"]]$Sample <- "PDX1233-tissue"
cap1[["PDX1264-cells"]]$Sample <- "PDX1264-cells"

capture1 <- cbind(
  cap1[["PDX1233-cells"]],
  cap1[["PDX1233-tissue"]],
  cap1[["PDX1264-cells"]]
)
```

Capture 2

```{r capture2}
cap2_samples <- c("PDX1233-cells", "PDX1233-tissue", "PDX1264-cells")
cap2_head <- "/stornext/Projects/score/GenomicsRnD/DB/S000379/R010_PDX_FLEX-2/outs/per_sample_outs/"
tail_path <- "/count/sample_filtered_feature_bc_matrix"

cap2 <- list()
for (i in cap2_samples) {
  path <- paste(cap2_head, i, tail_path, sep = "")
  cap2[[i]] <- read10xCounts(path, col.names = T)
}

cap2[["PDX1233-cells"]]$Sample <- "PDX1233-cells"
cap2[["PDX1233-tissue"]]$Sample <- "PDX1233-tissue"
cap2[["PDX1264-cells"]]$Sample <- "PDX1264-cells"

capture2 <- cbind(
  cap2[["PDX1233-cells"]],
  cap2[["PDX1233-tissue"]],
  cap2[["PDX1264-cells"]]
)
```

Combine both captures together

```{r}
dim(capture1)
dim(capture2)

capture1$Capture <- "One"
capture2$Capture <- "Two"

sce <- cbind(
  capture1,
  capture2
)
```

Add the gene symbols as row names

```{r}
row.names(sce) <- rowData(sce)$Symbol
```

Mitochondrial gene metrics and QC

```{r}
is.mito <- grepl("^MT", rownames(sce))
sce <- addPerCellQCMetrics(sce, subsets=list(Mito=is.mito))
sce <- addPerFeatureQC(sce)
```

Save outputs

```{r}
saveRDS(sce, "/stornext/Projects/score/Analyses/R010_multiplexing/SCEs/PDX/S000379_basic.sce.rds")
```

## Session Info

```{r}
sessionInfo()
```
