---
title: "Signal to noise plots"
author: "Patrick Grave and Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6,
                      fig.path='./figures/signalNoise/',
                      layout="l-body-outset")
knitr::opts_knit$set(root.dir = here::here())
suppressPackageStartupMessages({
  library(Matrix)
  library(tidyverse)
  library(ggridges)
  library(DropletUtils)
  library(patchwork)
})
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication(base_size = 24))
```

# Aim

In this notebook compare the oligo count distributions for the 3 protocols I tested.

## This sample
Run a more challenging sample = Ovarian carcinosarcome PDX nuclei.  
Four PDXs were processed with 10x Genomics nuclei isolation kit. A capture was immdiately performed which is the unlabelled.  
Three more aliquots were labelled in singlicate with cell multiplexing reagents and captured later.

## Comparisons

* Fresh unlabelled nuclei after nuclei isolation with 10x Genomics kit
* Fixed nuclei nuclei isolation with 10x Genomics kit
* Fresh nuclei CellPlex 1:10 diluted
* Fresh nuclei custom MULTI-Seq CMO (Nextera mod)
* Fresh nuclei 1ug nucleus hashtag Mab414

# Read data

These files are generated in scripts/R/S000322_write_tagCounts_file.R

```{r loadData}
CMO_matrix_dir <- here::here(
  "/stornext/Projects/score/GenomicsRnD/DB/S000322", "tag_counts", "Cellplex")

NXT_matrix_dir = here::here(
  "/stornext/Projects/score/GenomicsRnD/DB/S000322", "tag_counts", "MULTIseqCMO")

HTO_matrix_dir = here::here(
  "/stornext/Projects/score/GenomicsRnD/DB/S000322", "tag_counts", "HashtagAb")

matrix_dir <- list(CMO_matrix_dir, NXT_matrix_dir, HTO_matrix_dir)
names(matrix_dir) <- c("CMO", "NXT", "HTO")
```

# Read into tibble format {.tabset}

## CellPlex

```{r readCMO}
mat <- DropletUtils::read10xCounts(matrix_dir[["CMO"]])

cm <- as.data.frame(
  Matrix::as.matrix(
  t(counts(mat)
  )))
  
tb <- cm %>% 
  pivot_longer(cols = c(CMO301:CMO304),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )
tb$Protocol <- "CellPlex"
# Need to add 1 before log
tb$Count <- tb$Count + 1
tb_cmo <- tb
```

## MULTI-Seq CMO

```{r readLMO}
mat <-  DropletUtils::read10xCounts(matrix_dir[["NXT"]])
cm <- as.data.frame(
  Matrix::as.matrix(
  t(counts(mat)
  )))
  
tb <- cm %>% 
  pivot_longer(cols = c(Nxt_451:Nxt_455),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )
tb$Protocol <- "MULTI-Seq CMO"
# Need to add 1 before log
tb$Count <- tb$Count + 1

tb_lmo <- tb
```

## Hashtag Ab

```{r readNXT}
mat <-  DropletUtils::read10xCounts(matrix_dir[["HTO"]])
cm <- as.data.frame(
  Matrix::as.matrix(
  t(counts(mat)
  )))
  
tb <- cm %>% 
  pivot_longer(cols = c(HTO_1:HTO_4),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )
tb$Protocol <- "Hashtag Ab"
# Need to add 1 before log
tb$Count <- tb$Count + 1

tb_hto <- tb
```

We end by saving objects

```{r save}
tb <- list(
  tb_cmo,
  tb_lmo,
  tb_hto
)

names(tb) <- c("CellPlex", "MULTI-Seq CMO" ,"Hashtag Ab")
saveRDS(tb, here::here(
  "OvarianCarcinosarcoma/data/Matrices", "signal_noise_counts.tb.rds"
))
```

# GGridges plots {.tabset}

The plot I liked most is [https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html](GGridges)

## CellPlex

Looks better than on the cellranger summaries. Wider distribution than the other protocols indicative of comprimised cells.

```{r plt_cmo, fig.cap="CellPlex signal to noise"}
plt_cmo <- ggplot(tb_cmo, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges(alpha=0.66) +
  scale_x_continuous(trans='log10', limits = c(1,1e4)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 22) + 
  #scale_fill_brewer(palette = "Set1") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_cmo
```

## MULTI-Seq CMO

Bimodal distribution in most of the samples but not a big shift and the distributions merge into each other.

```{r plt_nxt, fig.cap="MULTI-Seq CMO signal to noise"}
plt_lmo <- ggplot(tb_lmo, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges(alpha=0.66) +
  scale_x_continuous(trans='log10', limits = c(1,1e4)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 22) + 
  #scale_fill_brewer(palette = "Dark2") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_lmo
```

## Hashtag Ab

Narrow distribution indicate less oligo swapping. However not a largee shift for positive cells.

```{r plt_hto, fig.cap="Hashtag Ab signal to noise"}
plt_hto <- ggplot(tb_hto, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges(alpha=0.66) +
  scale_x_continuous(trans='log10', limits = c(1,1e4)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 22) + 
  #scale_fill_brewer(palette = "Dark2") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_hto
```

# Gather ggridges into 1 plot

The aim is to reproduce the output of the cellranger report

```{r, fig.asp=1, fig.cap="Summary plot"}
plt_cmo / plt_lmo / plt_hto
```

# Visualise signal to noise

```{r}
sn <- read_csv(here::here(
  "OvarianCarcinosarcoma/data/Matrices", "signal_noise_ratio.csv"
))

sn$Protocol <- substr(sn$`CMO Name`, 1, 3)
sn$Protocol <- recode(sn$Protocol,
                      "HTO" = "Hashtag Ab",
                      "Nxt" = "MULTI-Seq CMO",
                      "CMO" = "CellPlex 1:10")

```

I copy pasted the signal to noise ratio from the cell rangere multi reports from each protocol.

Mung the names of the tags

```{r}
sn$CellRangerCall <- sn$`CMO Name`

sn$CellRangerCall <- recode(sn$CellRangerCall,
                      CMO301="Tag1",
                      CMO302="Tag2",
                      CMO303="Tag3",
                      CMO304="Tag4",
                      HTO_1="Tag1",
                      HTO_2="Tag2",
                      HTO_3= "Tag3",
                      HTO_4= "Tag4",
                      "Nxt_451"= "Tag1",
                      "Nxt_452"= "Tag2",
                      "Nxt_453"= "Tag3",
                      "Nxt_455"= "Tag4",
)
```


Something is different about the 4th sample. Maybe FACS sorting order which I need too check.  
Here the hashtag antibody has worked best although it is below the recommended signal to noise ratio of 5.

```{r signal_noise, fig.cap="Signal to noise from cell ranger"}
plt4 <- ggplot(data=sn) +
  geom_jitter(aes(y=`CMO signal-to-noise ratio`, x=Protocol, colour=CellRangerCall), 
              size=3, width = 0.15, show.legend=FALSE) +
  scale_y_continuous(breaks=seq(0,10,1)) +
  xlab("") + ylab("Signal to\n noise ratio")

plt4
```

## Session info
```{r sessionInfo}
sessionInfo()
```
