---
title: "Art of science 2023"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6,
                      fig.path='./figures/ArtOfScience/',
                      layout="l-body-outset")
knitr::opts_knit$set(root.dir = here::here())
library(here)
library(tidyverse)
library(ggalluvial)
library(knitr)
library(patchwork)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# Read data

Homotypic doublets are removed from this tibble

```{r}
tb <- read_csv("/stornext/Projects/score/Analyses/R010_multiplexing/Matrices/PBMC/vireo/pbmc_calls_rmHomoDoublets.csv.gz")

tb <- tb %>%
  select(cell_id, Protocol, vir_Donor, mul_Assignment, mul_Donor, chr_bff_cluster)

tb$Protocol <- recode(tb$Protocol,
                       HTO = "Hashtag Ab",
                       LMO = "MULTI-Seq",
                       CMO = "CellPlex")

homo_doublets <- read_csv("/stornext/Projects/score/Analyses/R010_multiplexing/Matrices/PBMC/vireo/pbmc_homo_doublets.csv")

remove_cells <- paste(homo_doublets$cell_barcode, homo_doublets$Protocol, sep = "_")

```


```{r}
ocm <- read_csv("/stornext/Projects/score/Analyses/R010_multiplexing/Matrices/PBMC/vireo/ocm/oca_cm.csv")

ocm$Protocol <- recode(ocm$Protocol,
                       HTO = "Hashtag Ab",
                       LMO = "MULTI-Seq",
                       CMO = "CellPlex")

ocm
```

## Sample classification specific accuracy

In case I need to show this data at some point.

```{r oca_cm}
sample_ocm <- read_csv("/stornext/Projects/score/Analyses/R010_multiplexing/Matrices/PBMC/vireo/ocm/sample_ca.csv")

sample_ocm$Protocol<- recode(sample_ocm$Protocol,
                       HTO = "Hashtag Ab",
                       LMO = "MULTI-Seq",
                       CMO = "CellPlex")

sample_ocm
```

# Plot as Sankey, alluvial plots or river plots {.tabset}

https://r-charts.com/flow/sankey-diagram-ggplot2/

The ggalluvial package provides the geom_alluvium and geom_stratum functions to create alluvial diagrams in ggplot2. You will need to pass your data in long format, specify the axis variables inside aes and use the previous functions.

```{r}
alluvial <- tb %>% 
  dplyr::count(vir_Donor, mul_Donor, mul_Assignment, Protocol, chr_bff_cluster) %>% 
  filter(vir_Donor != "unassigned") %>% 
  filter(chr_bff_cluster != "unassigned")

alluvial$vir_Donor <- recode(alluvial$vir_Donor,
  "VBDR1186" = "Sample1" ,
  "VBDR1206" = "Sample2",
  "VBDR1216" = "Sample3",
  "VBDR1233" = "Sample4"
  )

alluvial$mul_Donor <- recode(alluvial$mul_Donor,
  "VBDR1186" = "Sample1" ,
  "VBDR1206" = "Sample2",
  "VBDR1216" = "Sample3",
  "VBDR1233" = "Sample4"
  )

alluvial$chr_bff_cluster <- recode(alluvial$chr_bff_cluster,
  "CMO301" = "Tag1" ,
  "CMO302" = "Tag2",
  "CMO303" = "Tag3",
  "CMO304" = "Tag4",
  "CMO305" = "Tag5" ,
  "CMO306" = "Tag6",
  "CMO307" = "Tag7",
  "CMO308" = "Tag8"
  )
```

## Plot

```{r}
protocol <- "CellPlex"
current <- alluvial[alluvial$Protocol == protocol,]

current <- current %>% 
  arrange(mul_Donor)
```

```{r art_science}
am <- ggplot(data = current,
       aes(axis2=chr_bff_cluster,  axis3=mul_Donor, axis1 = vir_Donor,
           y = n)) +
  geom_alluvium(aes(fill = vir_Donor), discern=F) +
  geom_stratum(discern=F, alpha = .1, width = 1/3) +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum))) +
  scale_fill_brewer(palette = "Dark2") +
  theme_void() + theme(legend.position = "none")

am
```

### Session Info

```{r}
sessionInfo()
```
