---
title: "Heatmap of vireo and cell ranger multi agreement"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6,
                      fig.path='../figures/heat_vireo/',
                      layout="l-body-outset")
knitr::opts_knit$set(root.dir = here::here())
library(here)
library(tidyverse)
library(knitr)
library(pheatmap)
library(viridis)
library(patchwork)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

## Aim

Polished analysis to visualize data for manuscript.

# Read data

Homotypic doublets are retained in this tibble

```{r}
tb <- read_csv("/stornext/Projects/score/Analyses/R010_multiplexing/Matrices/PBMC/vireo/pbmc_calls_rmHomoDoublets.csv.gz")

tb <- tb %>%
  select(cell_id, Protocol, vir_Donor, mul_Assignment, mul_Donor, chr_bff_cluster)

tb$Protocol <- recode(tb$Protocol,
                       HTO = "Hashtag Ab",
                       LMO = "MULTI-Seq",
                       CMO = "CellPlex")

```

Recode the VBDR donors to generic labels

```{r}
tb$mul_Donor <- recode(tb$mul_Donor,
                      "VBDR1186" = "donor0",
                      "VBDR1206" = "donor1",
                      "VBDR1216" = "donor2",
                      "VBDR1233" = "donor3"
)

tb$vir_Donor <- recode(tb$vir_Donor,
                      "VBDR1186" = "donor0",
                      "VBDR1206" = "donor1",
                      "VBDR1216" = "donor2",
                      "VBDR1233" = "donor3"
)
```


```{r}
ocm <- read_csv("/stornext/Projects/score/Analyses/R010_multiplexing/Matrices/PBMC/vireo/ocm/oca_cm.csv")

ocm$Protocol <- recode(ocm$Protocol,
                       HTO = "Hashtag Ab",
                       LMO = "MULTI-Seq",
                       CMO = "CellPlex")

ocm$mul_Donor <- recode(ocm$mul_Donor,
                      "VBDR1186" = "donor0",
                      "VBDR1206" = "donor1",
                      "VBDR1216" = "donor2",
                      "VBDR1233" = "donor3"
)

ocm$vir_Donor <- recode(ocm$vir_Donor,
                      "VBDR1186" = "donor0",
                      "VBDR1206" = "donor1",
                      "VBDR1216" = "donor2",
                      "VBDR1233" = "donor3"
)

```

## Sample classification specific accuracy

In case I need to show this data at some point.

```{r oca_cm}
sample_ocm <- read_csv("/stornext/Projects/score/Analyses/R010_multiplexing/Matrices/PBMC/vireo/ocm/sample_ca.csv")

sample_ocm$Protocol<- recode(sample_ocm$Protocol,
                       HTO = "Hashtag Ab",
                       LMO = "MULTI-Seq",
                       CMO = "CellPlex")

sample_ocm$Classification <- recode(sample_ocm$Classification,
                      "VBDR1186" = "donor0",
                      "VBDR1206" = "donor1",
                      "VBDR1216" = "donor2",
                      "VBDR1233" = "donor3"
)

sample_ocm$vir_Donor <- recode(sample_ocm$vir_Donor,
                      "VBDR1186" = "donor0",
                      "VBDR1206" = "donor1",
                      "VBDR1216" = "donor2",
                      "VBDR1233" = "donor3"
)

sample_ocm$n <- as.integer(sample_ocm$n)
```

# Generate heatmaps {.tabset}

```{r}
font_size <- 16
fontsizeNumber <- 16
```

## CellPlex

Mung into shape

```{r}
protocol <- "CellPlex"

crm <- sample_ocm %>% 
  filter(Protocol == protocol) %>% 
  filter(algorithm == "Cell Ranger") %>% 
  select(vir_Donor, Classification,Protocol, n) %>% 
  pivot_wider(names_from = vir_Donor, values_from = n)

crm <- as.data.frame(crm)
row.names(crm) <- crm$Classification
crm <- as.matrix(crm[,c(3:8)])
# Recode NA to 0
crm[is.na(crm)] <- 0
```

Generate the heatmap

```{r cellplex_heat}
pheatmap(crm,
         cluster_rows=F,
         cluster_cols = F,
         treeheight_col=0,
         treeheight_row=0,
         show_rownames=TRUE,
         show_colnames = TRUE,
         display_numbers = TRUE,
         color = viridis(100), 
         number_color = "white",
         fontsize=font_size, 
         number_format = "%0.f", 
         fontsize_number = fontsizeNumber,
         main="CellPlex"
         )
```

## Hashtag Ab

Mung into shape

```{r}
protocol <- "Hashtag Ab"

crm <- sample_ocm %>% 
  filter(Protocol == protocol) %>% 
  filter(algorithm == "Cell Ranger") %>% 
  select(vir_Donor, Classification,Protocol, n) %>% 
  pivot_wider(names_from = vir_Donor, values_from = n)

crm <- as.data.frame(crm)
row.names(crm) <- crm$Classification
crm <- as.matrix(crm[,c(3:8)])
# Recode NA to 0
crm[is.na(crm)] <- 0
```

Generate the heatmap

```{r hashtag_heat}
pheatmap(crm,
         cluster_rows=F,
         cluster_cols = F,
         treeheight_col=0,
         treeheight_row=0,
         show_rownames=TRUE,
         show_colnames = TRUE,
         display_numbers = TRUE,
         color = viridis(100), 
         number_color = "white",
         fontsize=font_size, 
         number_format = "%0.f", 
         fontsize_number = fontsizeNumber,
         main="Hashtag Ab"
         )
```

## MULTI-Seq LMO

Mung into shape

```{r}
protocol <- "MULTI-Seq"

crm <- sample_ocm %>% 
  filter(Protocol == protocol) %>% 
  filter(algorithm == "Cell Ranger") %>% 
  select(vir_Donor, Classification,Protocol, n) %>% 
  pivot_wider(names_from = vir_Donor, values_from = n)

crm <- as.data.frame(crm)
row.names(crm) <- crm$Classification
crm <- as.matrix(crm[,c(3:8)])
# Recode NA to 0
crm[is.na(crm)] <- 0
```

Generate the heatmap

```{r multiseq_heat}
pheatmap(crm,
         cluster_rows=F,
         cluster_cols = F,
         treeheight_col=0,
         treeheight_row=0,
         show_rownames=TRUE,
         show_colnames = TRUE,
         display_numbers = TRUE,
         color = viridis(100), 
         number_color = "white",
         fontsize=font_size, 
         number_format = "%0.f", 
         fontsize_number = fontsizeNumber,
         main="MULTI-Seq"
         )
```

### Session Info

```{r}
sessionInfo()
```
