---
title: "Create Seurat object"
author: "Daniel Brown"
description: "Quality control and normalise"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6, 
                      fig.path='./figures/create_SRT/',
                      warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(Seurat)
library(here)
library(knitr)
library(tidyverse)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# Aim

I downloaded the data from 

# https://singlecell.broadinstitute.org/single_cell/study/SCP1290/molecular-logic-of-cellular-diversification-in-the-mammalian-cerebral-cortex#/

To build a reference to map to with Seurat Mapping and annotating query datasets vignette.
https://satijalab.org/seurat/articles/integration_mapping.html

# Read data

## Metadata

```{r}
meta <- read_tsv(here::here(
  "data/SCP1290_mouseEmbryoBrainRegev", "metaData_scDevSC.txt"
))

meta <- meta[c(2:98048),]

umap <- read.delim(here::here(
  "data/SCP1290_mouseEmbryoBrainRegev", "cluster_scDevSC.merged.umap.txt"
))
umap <- umap[c(2:98048),]

tsne <- read_tsv(here::here(
  "data/SCP1290_mouseEmbryoBrainRegev", "cluster_scDevSC.merged.tsne.txt"
))
tsne <- tsne[c(2:98048),]
```

Filter the metadata for E18 cells which is the same timepoint as my study.

```{r}
keep_cells <- grepl("^E18", meta$NAME)

meta <- meta[keep_cells,]
umap <- umap[keep_cells,]
tsne <- tsne[keep_cells,]
```

meta object needs recoded for column count write to disk and re read.

```{r}
write_csv(meta, here::here(
  "data/SCP1290_mouseEmbryoBrainRegev", "metaData_scDevSC.csv.gz"
))

meta <- read_csv(here::here(
  "data/SCP1290_mouseEmbryoBrainRegev", "metaData_scDevSC.csv.gz"
))
meta <- as.data.frame(meta)
row.names(meta) <- meta$NAME
```


## Seurat object

```{r loadData}
counts <- Read10X(here::here(
  "data/SCP1290_mouseEmbryoBrainRegev"))

counts_filter <- counts[,keep_cells]

dim(counts_filter)
```

This object contains all the cells whereas I need only cells from E18.5 embryo.

```{r}
srt <- CreateSeuratObject(
  counts = counts_filter, 
  project = "SCP1290",
  meta.data = meta
)
```

# Examine the metadata fields

The broad cell type labels

```{r}
unique(srt$Gral_cellType)
```

The finer labels

```{r}
unique(srt$New_cellType)
```

# Add the UMAP coordinates

Need to mung the umap into a dimred object.

```{r}
umap$X <- as.double(umap$X)
umap$Y <- as.double(umap$Y)
row.names(umap) <- umap$NAME
umap <- umap[,c(2,3)]
umap <- as.matrix(umap)
colnames(umap) <- c("umap_pub1", "umap_pub2")

umap1 <- umap[colnames(srt),]

umap1 <- CreateDimReducObject(embeddings = umap1, key = "umap_pub")
srt[["umap_pub"]] <- umap1
```

## Filter seurat object

Remove doublets

```{r}
srt <- srt[,!srt$Doublet_intersect == "Doublet"]

DimPlot(object = srt, reduction = "umap", group.by = "Gral_cellType")
```

## Normalise

```{r}
srt <- SCTransform(srt, verbose = TRUE)
```

Perform dimensionality reduction by PCA and UMAP embedding

```{r}
srt <- RunPCA(srt, verbose = T)
srt <- RunUMAP(srt, dims = 1:30, verbose = T)

srt <- FindNeighbors(srt, dims = 1:30, verbose = T)
srt <- FindClusters(srt, verbose = T)

p1 <- DimPlot(srt, label = TRUE, group.by = "Gral_cellType", reduction = "umap") + NoLegend()
p2 <- DimPlot(srt, label = TRUE, group.by = "Gral_cellType", reduction = "umap_pub") + NoLegend()

p1 + p2
```

# Save object

```{r}
saveRDS(srt, here::here(
  "data/SCP1290_mouseEmbryoBrainRegev", "SCP1290_mouseEmbryo.srt.rds"
))
```


## Session Info

```{r}
sessionInfo()
```
