---
title: "CRISPR clean offtarget effects"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6, 
                      fig.path='./figures/off_target/',
                      warning=FALSE, message=FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(scater)
library(scran)
library(patchwork)
library(ggvenn)
library(tidyverse)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# Aim

Evaluate jumpcode CRISPRclean single-cell boost kit. This adds more content on top of ribosomal and mitochondrial genes.  
Including Pseudogenes and non variablee genes.

In this notebook perform the DE testing between treated and untreated.

## This sample

I restrict my analysis to R010_MULTI-Seq_GEX in PBMCs. This sample had the best signal to noise.

I sequenced this particular sample much deeper as part of NN265.  
[ELN entry](https://au-mynotebook.labarchives.com/share/Daniel_RnD_Projects/MzQ3LjF8NjgxOTcvMjY3LTE0NzgvVHJlZU5vZGUvMjMwMTMwNzYyN3w4ODEuMDk5OTk5OTk5OTk5OQ==)

## Comparisons

I mapped both the undepleted and CRISPRclean depleted datasets to GRCh38 reference provided by cellranger and the masked reference provided by jumpcode. Jumpcode says only the masked reference will highlight the differences in depleted.

## JUMPcode scCLEAN publication

The folloowing quotes from the preprint

*Pandey, Amitabh C., Jon Bezney, Dante DeAscanis, Ethan Kirsch, Farin Ahmed, Austin Crinklaw, Kumari Sonal Choudhary et al. "A CRISPR/Cas9-based enhancement of high-throughput single-cell transcriptomics." bioRxiv (2022): 2022-09.*

* While 33 genes passed the significance threshold, only 21 were determined to be off-target and 18 were direct targets of scCLEAN guides (Supplementary Fig. 1d-e).
* Distinguishing between genes that are a product of random sampling from two libraries of fundamentally different composition and true off-target effects. 
  + Genes that don’t follow a linear trend in log2 fold change reduction (negative values represent a decrease in scCLEAN expression relative to 10x-V3) are considered offtarget (n=21) and are automatically detected by a knee bend algorithm. 
* e, Quantifying whether the 21 selected genes in this hierarchical selection process are targets of scCLEAN guides or overlap with regions of on-target genes (18/21)

# Load Data

Read in data

```{r loadData, eval=FALSE}
sces <- readRDS(here::here(
  "CRISPRclean/data/Jumpcode/SCEs", "PBMC_CRISPR_colData.sces.rds"
))
names(sces)

sce_jmp <- sces[["Jumpcode_full"]]
sce_utd <- sces[["Untreated"]]

sce <- cbind(sce_jmp, sce_utd)
rm(sces, sce_utd, sce_jmp)
```

Read the differential expressed genes between Jumpcode and untreated from *3B_DEtest_crisprClean* notebook.  
Subset for the significant off-target elements.

```{r}
tb <- read_csv(here::here(
  "CRISPRclean/data/Jumpcode/Matrices", "PBMC_DEgenes.csv.gz"
))

db_expt <- tb %>% 
  filter(logFC >= 1 & PValue < 0.05) %>% 
  filter(In_Panel == FALSE) %>% 
  select(DElabel)

db_expt <- db_expt$DElabel
db_expt <- db_expt[!is.na(db_expt)]
```

Remove the novel transcripts which are unannotated genes, pseduogenes, miRNAs, lncRNAs etc They start with "AC" pr "AL"

```{r}
db_expt <- db_expt[grep("^AC" ,db_expt, invert = TRUE)]
db_expt <- db_expt[grep("^AL" ,db_expt, invert = TRUE)]
```

Load the jumpcode panel

```{r}
jpc <- read_csv(here::here(
  "CRISPRclean/metadata/Jumpcode", "CRISPRclean_Single_Cell_RNA_Boost_Kit_depletion_content_set.csv"
))
```

Load the off target panel from 

```{r}
offtarget <- read_csv(here::here(
  "CRISPRclean/metadata/Jumpcode", "Paandey_2022_offtarget.csv"
))
```

# Create venn digram

```{r mung_venn, fig.dim=c(10,14)}
jpc_expt <- offtarget$GeneID

venn_list <- list(
  "scBoost Brown" = db_expt, 
  "Pandey2022" = jpc_expt
  )

ggvenn(venn_list,
       show_elements = T, label_sep = "\n",
       fill_color = c("blue", "green")
       )

```

# Conclusion

There is some overlap between the list produced by Pandey and my DE list but not 100%.  
Without replicates hard to make much comment on the genes I find t be differentially expressed outside the Pandey list.

## Session Info

```{r}
sessionInfo()
```
