---
title: "Signal to noise plots"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6,
                      fig.path='./figures/signalNoise/',
                      layout="l-body-outset")
knitr::opts_knit$set(root.dir = here::here())
library(Matrix)
library(tidyverse)
library(cowplot)
library(GGally)
library(ggridges)
library(DropletUtils)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

## Aim

I downloaded this dataset from 10x Genomics datasets. It is v3.1 chemistry processed with Cellranger v6.0.
I want to compare to the dataset I generated for CellPlex.

# Read data

These files are generated in scripts/R/NN236_write_tagCounts_file.R

```{r loadData}
sce <- readRDS(here::here(
  "data/10xPBMC_2CMOs", "10x-basic.rds"))
```

## Read into tibble format

```{r matrix, cache=TRUE}
mat <- Matrix::as.matrix(
  counts(altExp(sce
                )))

mat <- as.data.frame(t(mat))
  
tb <- mat %>% 
  pivot_longer(cols = c(CMO301:CMO308),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )

tb$Protocol <- "CMO"
# Need to add 1 before log
tb$Count <- tb$Count + 1
tb_cmo <- tb
```

# GGridges

The plot I liked most is [https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html](GGridges)

CellPlex had the highest background.

```{r plt_cmo, fig.cap="CellPlex signal to noise"}
plt_cmo <- ggplot(tb, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges() +
  scale_x_continuous(trans='log10', limits = c(1,1e5)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 18) + 
  scale_fill_brewer(palette = "Dark2") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_cmo
```

# Pairwise plots with GGally {.tabset}

Here we reproduce the pair-wise plots from Cellranger MULTI.
I find these plots are too small to be particularly visible.

```{r cmo_scatter, fig.cap="Note the red dots denote all doublets so do appear in the right and top part of the graph"}
my_colors <- c("gold", "darkgreen", "blue", 
               "red", "black","orange")


my_sample_col <- as.data.frame(sce$mul_Assignment)
colnames(my_sample_col) <- "mul_Assignment"

c12 <- cbind(
  mat[,c("CMO301", "CMO302")],
  my_sample_col)

# Recode the other tags to singlets
# Otherwise there are too many colours plotted
c12$mul_Assignment <- recode(c12$mul_Assignment,
                            CMO304 = "Other Singlet",
                            CMO305 = "Other Singlet",
                            CMO306 = "Other Singlet",
                            CMO307 = "Other Singlet",
                            CMO308 = "Other Singlet",
                            CMO303 = "Other Singlet",
                            CMO309 = "Other Singlet",
                            CMO310 = "Other Singlet",
                            CMO311 = "Other Singlet",
                            CMO312 = "Other Singlet")

p1 <- ggplot(c12, 
             aes(x = CMO301, y= CMO302, colour = mul_Assignment)) + 
  geom_point(size=1, alpha=0.25) + 
  ylab("Barcode 2") + 
  xlab("Barcode 1") +
  ggtitle("CellPlex") +
  scale_colour_manual(values=my_colors) +
  scale_y_continuous(trans='log10', limits = c(NA,100000)) +
  scale_x_continuous(trans='log10', limits = c(NA,100000)) +
  annotation_logticks(base = 10, sides = "bl")

p1
```

# Conclusion

The 10x Genomics dataset is much cleaner than my one. Some differences:

* No superloading 13,446 cells out
* Only 2 tags used

## Session info
```{r sessionInfo}
sessionInfo()
```
