---
title: "Titrate MULTI-Seq reagent by FACS"
description: |
  Titrate CellPlex reagent
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.asp=0.618, 
                      fig.path='./Figures/MULTISeq/',
                      warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
library(flowCore)
library(ggcyto)
library(knitr)
library(tidyverse)
source(here::here("scripts/R/theme_publication.R"))
```

# Installation
Follow the guidelines of 
https://jchellmuth.com/posts/FACS-with-R/

```{r, eval=FALSE}
BiocManager::install(c("flowCore", "ggcyto"))
```

# Experimental details

## Recap
The first test of MULTI-seq reagents by FACS was a success. However the titration was less than 10 fold probably due to the nature of hybridizing an anchor and co-anchor oligo.

## Samples
1.	K562 cells

## Aim

Titrate MULTI-Seq oligos

# Read and format .fcs data
The read.flowSet allows you to read many .fcs at once. The resultant flowSet object will store the data from all these .fcs files.

```{r, include=TRUE}
list.files(here::here(
  "data/FACS/", "210929_check_customMULTIseq"
  ))

fs <- read.flowSet(path = here::here(
  "data/FACS/", "191030_MULTIseq_FACS"
  ),
  pattern = "*.fcs",
  alter.names = T)

filenames <- as_tibble(pData(fs))
```

Sample information (aka phenotypic data) such as sample names can be accessed with the pData function. The default sample names are not very useful. So, let’s extract the sample name.

## Mung sample names

Daniel had to make this samplesheet himself. Not sure what the correct controls are.

```{r}
samples <- read_csv(here::here(
  "metadata/FACS", "191030_multiSeq_titr.csv"
))

#pData(fs)$label <- samples$label
pData(fs)$concentration <- samples$concentration
pData(fs)$Oligo <- samples$Oligo
pData(fs)$Detection <- samples$Detection
pData(fs)$sample <- paste(samples$Oligo,
                          "detect",
                          samples$Detection,
                          samples$concentration,
                          "dil",
                          sep="-")

pData(fs) %>% 
  kable()
```

```{r}
colnames(fs)[colnames(fs)=="R660.20.A"] <- "Feature BC 2"
colnames(fs)[colnames(fs)=="Alexa.594.A"] <- "Feature BC 1"
colnames(fs)[colnames(fs)=="V450.50.A"] <- "DAPI"

colnames(fs) %>% 
  kable()
```

## Convert to Gating set object
To be able to add gates, the flowSet has to be transformed to a GatingSet object with the GatingSet function

```{r}
gs <- GatingSet(fs)
```

If you want to check where the singlets you filtered out show up in FSC vs SSC (the plot that is typically used to remove debris or gate live cells), you can do this by setting different subets in the ggcyto command (‘root’ or ‘Singlets’).

# Gating strategy {.tabset}

## Debris gate

The axis limits are set by the instrument. Use FSC.H,y=SSC.H

```{r}
ggcyto(gs[[1]], 
       aes(x=FSC.H,y=SSC.A), subset="root") + 
  geom_hex(bins = 200) + 
  ggcyto_par_set(limits = "instrument") + 
    scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Define the gate
check the gate in a plot (before adding it to the gating set)
if the gate seems right, add / apply it to the gating set
recompute the gatingSet to have statistics available.

Its much easier to draw rectangle gates than polygon gates in R.

```{r}
# define gate
g.debris <- rectangleGate(filterId = "noDebris",
                          "FSC.H"=c(2.2e4,1.2e5),
                          "SSC.A"=c(2.5e4,1.5e5)) 

# check gate
ggcyto(gs[[12]],aes(x=FSC.H,y=SSC.A) ,subset="root") + 
  geom_hex(bins = 200) + 
  geom_gate(g.debris) + 
  ggcyto_par_set(limits = "instrument") + 
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gate the cells

```{r}
gs_pop_add(gs,g.debris) # add gate to GatingSet
recompute(gs)
```

```{r debris_gate, fig.height=20, fig.width=12, fig.cap = "Debris gate"}
ggcyto(gs,aes(x=FSC.H,y=SSC.A), 
       subset="noDebris") + 
  geom_hex(bins = 100) + 
  geom_gate("noDebris") +
  geom_stats(adjust = 0.8) + 
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) + facet_wrap(~sample, ncol = 3)
```

## Singlet gate

Drawing the polygon gate is a real pain, I have to manually type numbers in.

### Define the gate
check the gate in a plot (before adding it to the gating set)
if the gate seems right, add / apply it to the gating set
recompute the gatingSet to have statistics available

```{r}
# define gate
g.singlets <- polygonGate(filterId = "Singlets",
                           "FSC.A"=c(4e4,2e5,8e4,3e4),
                           "FSC.H"=c(1.5e4,1e5,1e5,2e4)) 

# check gate
ggcyto(gs[[9]],aes(x=FSC.A,y=FSC.H) ,subset="noDebris") + 
  geom_hex(bins = 200) + 
  geom_gate(g.singlets) + 
  ggcyto_par_set(limits = "instrument") + 
      scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gate the cells

```{r}
gs_pop_add(gs,g.singlets,
           parent="noDebris") # add gate to GatingSet

recompute(gs)
```

overview plots at each gating step.

```{r singlet_gate, fig.height=20, fig.width=12, fig.cap = "Singlet gate"}
ggcyto(gs,aes(x=FSC.A,y=FSC.H), 
       subset="noDebris") + 
  geom_hex(bins = 200) + 
  geom_gate("Singlets") +
  geom_stats(adjust = 0.8) + 
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) + facet_wrap(~sample, ncol = 5)
```

## Live cell gate

I'm unsure if PI or DAPI is the viability dye. Use DAPI because spectral overlap means PI is useless without compensation.

Gating steps are as above. Differentiating live and dead cells solely based on DAPI exclusion.

```{r}
g.live <- rectangleGate(filterId = "Live",
                      "FSC.A"=c(4e4,1.8e5),
                      "DAPI"=c(6,8e2)) # define gate

# Add biexponential axis
ggcyto(gs[[1]], aes(x=FSC.A,y=DAPI),
       subset="Singlets") + 
  geom_hex(bins = 200) + 
  geom_gate(g.live) + 
  scale_y_flowjo_biexp() +
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gating out dead cells

```{r}
gs_pop_add(gs, 
    g.live, 
    parent="Singlets") # add gate to GatingSet

recompute(gs)
```

```{r live_gate, fig.height=20, fig.width=12, fig.cap = "Live gate"}
ggcyto(gs,aes(x=FSC.A,y=DAPI),
       subset="Singlets") + 
  geom_hex(bins = 200) + 
  geom_gate("Live") +
  geom_stats(adjust = 0.8) + 
  scale_y_flowjo_biexp() +
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) +  facet_wrap(~sample, ncol = 4)
```

# Set Alexa 647+ feature barcode 2 detection oligo

This looks like the best control to use to set the gate.
* Unstained PI sample

```{r}
g.a647 <- rectangleGate(filterId="Feature BC 2",
                       "Feature BC 2"=c(3e2, Inf)) # set gate

ggcyto(gs[[4]],
       aes(x=`Feature BC 2`), subset="Live") + 
  geom_density(fill="rosybrown") + 
  geom_gate(g.a647) + 
  ggcyto_par_set(limits = "instrument") + 
  scale_x_flowjo_biexp() + 
    scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 12))
```

## Add gate

```{r}
gs_pop_add(gs, g.a647, parent="Live") # add gate to GatingSet
recompute(gs) # recalculate Gatingset

ggcyto(gs, aes(x=`Feature BC 2`),
       subset="Live") + 
  geom_density(fill="rosybrown") + 
  geom_gate("Feature BC 2") +
  geom_stats(adjust = 0.99, digits = 3) +
  ggcyto_par_set(limits = "instrument") + 
  scale_x_flowjo_biexp() +
  xlab("Feature barcode detection oligo") +
  facet_wrap(~sample , ncol = 4) +
  scale_colour_Publication()+ theme_Publication() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1), strip.text.x = element_text(size = 8)) +
  facet_wrap(~sample)
```

# Visualise Alexa 647 signal {.tabset}

## By Dilution

```{r, include=FALSE}
pData(gs)$sample
# Subset only the relevant columns

gs_subset <- gs[c(4,6:12),]
pData(gs_subset)$concentration <- as.factor(pData(gs_subset)$concentration)
```

Make overlay plot

```{r FB2_dil_gate, fig.cap = "Alexa 647 gate by dilution"}  
ggcyto(gs_subset, aes(x=`Feature BC 2`, fill=concentration), 
       subset="Live",) + 
  geom_density(alpha=0.33) + 
  geom_gate("Feature BC 2") +
  ggcyto_par_set(limits = "instrument") + 
  scale_x_flowjo_biexp() +
  facet_grid(~Oligo) +
  xlab("Feature barcode detection oligo") +
  ggtitle("") +
  scale_fill_discrete(limits=c("1", "2", "4", "8", "32", "128", "0")) +  
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 12))
```

### Cleanup
```{r, eval=F}
write_csv(psm, here::here(
  "data/FACS/191030_multiseq_dil", "population_stats.csv"
  ), quote = "none")

sessionInfo()
```