---
title: "Compare pooled washing MULTI-Seq"
description: |
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.asp=0.6,
                      dev = "png", 
                      dpi = 300,
                      fig.path='./Figures/pooledWashing/',
                      warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
library(flowCore)
library(ggcyto)
library(knitr)
library(tidyverse)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# Installation
Follow the guidelines of 
https://jchellmuth.com/posts/FACS-with-R/

```{r, eval=FALSE}
BiocManager::install(c("flowCore", "ggcyto"))
```

# Experimental details

For the cell multiplexing manuscript I want to compare pooled and separate washing for LMO and CMO labeling.

Additionally I have been using MULTI-Seq with feature barcode 2 capture sequence. I need to compare to the original poly-A capture sequence as supplemental data.

* Lipid oligos
+ 50uM MULTI-Seq LMO from Chris McGinnis
* LMO
+ Custom CMO from IDT
* CMO

Barcoding oligos

* 100uM MULTI-Seq FB2 BC2 = Gin
* 100uM Nxt_multi_451 = Nxt
* 100uM Tom’s poly A MULTI-Seq = poly-A

Detection oligos

* Feature barcode detection oligo A647
* Feature barcode detection oligo A549
* Poly-T A647

Prepare a single cell suspension of K562 cells.

# Read and format .fcs data
The read.flowSet allows you to read many .fcs at once. The resultant flowSet object will store the data from all these .fcs files.

```{r, include=TRUE}
list.files(here::here(
  "data/FACS/", "211110_MULTIseq_compare"
  ))

fs <- read.flowSet(path = here::here(
  "data/FACS/", "211110_MULTIseq_compare"
  ),
  pattern = "*.fcs",
  alter.names = T)

filenames <- as_tibble(pData(fs))
```

Sample information (aka phenotypic data) such as sample names can be accessed with the pData function. The default sample names are not very useful. So, let’s extract the sample name.

## Mung sample names

Daniel had to make this samplesheet himself.

```{r, include=FALSE}
samples <- read_csv(here::here(
  "metadata/FACS", "211110_MULTIseq_compare_samplesheet.csv"
))

# Subset and the relevant rows for this experiment
samples <- samples[c(1,2,8:24),]
samples$name
filenames$name

#pData(fs)$label <- samples$label
pData(fs)$barcode <- samples$Barcode
pData(fs)$detection <- samples$Detection
pData(fs)$replicate <- samples$replicate
pData(fs)$label <- samples$label

pData(fs)$sample <- paste(samples$label,
                          samples$Barcode,
                          samples$Detection,
                          "rep",
                          samples$replicate,
                          sep="-")
```

Write phenotypic data

```{r}
pData(fs) %>% 
  kable()
```

Rename channels.

```{r}
colnames(fs)[colnames(fs)=="APC.A"] <- "A647"
colnames(fs)[colnames(fs)=="Alexa.594.A"] <- "A594"
```

## Convert to Gating set object
To be able to add gates, the flowSet has to be transformed to a GatingSet object with the GatingSet function

```{r}
gs <- GatingSet(fs)
```

If you want to check where the singlets you filtered out show up in FSC vs SSC (the plot that is typically used to remove debris or gate live cells), you can do this by setting different subets in the ggcyto command (‘root’ or ‘Singlets’).

# Gating strategy {.tabset}

## Debris gate

The axis limits are set by the instrument. Use FSC.H,y=SSC.H

```{r}
ggcyto(gs[[1]], 
       aes(x=FSC.H,y=SSC.A), subset="root") + 
  geom_hex(bins = 200) + 
  ggcyto_par_set(limits = "instrument") + 
    scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Define the gate
check the gate in a plot (before adding it to the gating set)
if the gate seems right, add / apply it to the gating set
recompute the gatingSet to have statistics available.

Its much easier to draw rectangle gates than polygon gates in R.

```{r}
# define gate
g.debris <- rectangleGate(filterId = "noDebris",
                          "FSC.H"=c(2.2e4,8e4),
                          "SSC.A"=c(2.75e4,1.5e5)) 

# check gate
ggcyto(gs[[1]],aes(x=FSC.H,y=SSC.A) ,subset="root") + 
  geom_hex(bins = 200) + 
  geom_gate(g.debris) + 
  ggcyto_par_set(limits = "instrument") + 
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gate the cells

```{r}
gs_pop_add(gs,g.debris) # add gate to GatingSet
recompute(gs)
```

```{r debris_gate, fig.height=20, fig.width=12, fig.cap = "Debris gate"}
ggcyto(gs,aes(x=FSC.H,y=SSC.A)) +
  geom_hex(bins = 200) + 
  geom_gate("noDebris") +
  geom_stats(adjust = 0.8) + 
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) + facet_wrap(~sample, ncol = 3)
```

## Singlet gate

Drawing the polygon gate is a real pain, I have to manually type numbers in.

### Define the gate
check the gate in a plot (before adding it to the gating set)
if the gate seems right, add / apply it to the gating set
recompute the gatingSet to have statistics available

```{r}
# define gate
g.singlets <- polygonGate(filterId = "Singlets",
                           "FSC.A"=c(4e4,1.8e5,8e4,3e4),
                           "FSC.H"=c(1.5e4,8e4,9e4,2e4)) 

# check gate
ggcyto(gs[[1]],aes(x=FSC.A,y=FSC.H) ,subset="noDebris") + 
  geom_hex(bins = 200) + 
  geom_gate(g.singlets) + 
  ggcyto_par_set(limits = "instrument") + 
      scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gate the cells

```{r}
gs_pop_add(gs,g.singlets,
           parent="noDebris") # add gate to GatingSet

recompute(gs)
```

overview plots at each gating step.

```{r singlet_gate, fig.height=8, fig.width=12, fig.cap = "Singlet gate"}
ggcyto(gs,aes(x=FSC.A,y=FSC.H), 
       subset="noDebris") + 
  geom_hex(bins = 200) + 
  geom_gate("Singlets") +
  geom_stats(adjust = 0.8) + 
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) + facet_wrap(~sample, ncol = 5)
```

## Live cell gate

I'm unsure if PI or DAPI is the viability dye. Use DAPI because spectral overlap means PI is useless without compensation.

Gating steps are as above. Differentiating live and dead cells solely based on DAPI exclusion.

```{r}
g.live <- rectangleGate(filterId = "Live",
                      "FSC.A"=c(5e4,1.5e5),
                      "DAPI.A"=c(100,9e2)) # define gate

# Add biexponential axis
ggcyto(gs[[1]], aes(x=FSC.A,y=DAPI.A),
       subset="Singlets") + 
  geom_hex(bins = 200) + 
  geom_gate(g.live) + 
  scale_y_flowjo_biexp() +
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gating out dead cells

```{r}
gs_pop_add(gs, 
    g.live, 
    parent="Singlets") # add gate to GatingSet

recompute(gs)
```

```{r live_gate, fig.height=20, fig.width=12, fig.cap = "Live gate"}
ggcyto(gs,aes(x=FSC.A,y=DAPI.A),
       subset="Singlets") + 
  geom_hex(bins = 200) + 
  geom_gate("Live") +
  geom_stats(adjust = 0.8) + 
  scale_y_flowjo_biexp() +
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) +  facet_wrap(~sample, ncol = 4)
```

# Set double labelled gate

Extract only those relevant samples

```{r}
pData(gs)$sample

#gs_orig <- gs
gs <- gs[c(5:8,13,14,17,18),]
```


```{r, eval=T}
g.double <- rectangleGate(filterId="Barcode_Pos",
               "A647"=c(2.5e3,1e5), 
               "A594"=c(1.5e3,1e5))


# check gate
ggcyto(gs[[5]],aes(x=A647,y=A594) ,subset="Live") + 
  geom_hex(bins = 200) + 
  geom_gate(g.double) + 
  ggcyto_par_set(limits = "instrument") + 
  scale_x_flowjo_biexp() + scale_y_flowjo_biexp() +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

```{r}
gs_pop_add(gs, 
    g.double, 
    parent="Live") # add gate to GatingSet

recompute(gs)
```

```{r fig.height=8, fig.width=12}
ggcyto(gs,
       aes(x=A647, y=A594), subset="Live") + 
  geom_hex(bins = 200) + 
  geom_gate(g.double) + 
  geom_stats(adjust = 0.8) + 
  ggcyto_par_set(limits = "instrument") + 
  scale_x_flowjo_biexp() + scale_y_flowjo_biexp() +
  theme_Publication() +
  facet_wrap(~sample)
```

### Cleanup
```{r}
sessionInfo()
```