---
title: "Compare MULTI-Seq LMO vendors"
description: |
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.asp=0.6,
                      dev = "png", 
                      dpi = 300,
                      fig.path='./Figures/comp_MULTIseqVendors/',
                      warning=FALSE, message=FALSE)
```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
library(flowCore)
library(ggcyto)
library(knitr)
library(tidyverse)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# Installation
Follow the guidelines of 
https://jchellmuth.com/posts/FACS-with-R/

```{r, eval=FALSE}
BiocManager::install(c("flowCore", "ggcyto"))
```

# Experimental details

## Aim
I have been using MULTI-Seq LMO oligos from Chris McGinnis. This worked well in a comparison experiment on PBMCs but is a finite resource. 
I received MULTI-Seq LMOs from Sigma Merck under MTA and will test them by FACS before a scRNA-Seq experiment. Additionally check the 2nd batch of Genelink LMO I was sent.

Secondary aim: play with BlueWasher settings to get good cell pellet and low supernatant.
Reagents to test:

1.	50uM MULTI-Seq LMO from Chris McGinnis
  +	MGN LMO
2.	MULTI-Seq from Sigma Merck
  +	SIG LMO
3.	MULTI-Seq LMO from Genelink
  +	GLK LMO
  + Supplement with 20% DMSO in TE

Conduct the experiment in triplicate FACS tubes = 9 tubes.
Samples
Prepare a single-cell suspension of K562 cells a minimum of 100,000 cells per well.

# Read and format .fcs data
The read.flowSet allows you to read many .fcs at once. The resultant flowSet object will store the data from all these .fcs files.

```{r, include=TRUE}
list.files(here::here(
  "data/FACS/", "220520_MULTIseq_compareVendor"
  ))

fs <- read.flowSet(path = here::here(
  "data/FACS/", "220520_MULTIseq_compareVendor"
  ),
  pattern = "*.fcs",
  alter.names = T)

filenames <- as_tibble(pData(fs))
```

Sample information (aka phenotypic data) such as sample names can be accessed with the pData function. The default sample names are not very useful. So, let’s extract the sample name.

## Mung sample names

Daniel had to make this samplesheet himself.

```{r, include=FALSE}
samples <- read_csv(here::here(
  "metadata/FACS", "220523_MULTIseq_vendor_samplesheet.csv"
))

pData(fs)$Vendor <- samples$Vendor
pData(fs)$Detection <- samples$Detection
pData(fs)$replicate <- samples$replicate

pData(fs)$sample <- paste(samples$Vendor,
                          samples$replicate,
                          sep="-")
```

Write phenotypic data

```{r}
pData(fs) %>% 
  select(-name) %>% 
  kable()
```

Rename channels.

```{r}
colnames(fs)[colnames(fs)=="APC.A"] <- "A647"
colnames(fs)[colnames(fs)=="Alexa.594.A"] <- "A594"
```

## Convert to Gating set object
To be able to add gates, the flowSet has to be transformed to a GatingSet object with the GatingSet function

```{r}
gs <- GatingSet(fs)
```

If you want to check where the singlets you filtered out show up in FSC vs SSC (the plot that is typically used to remove debris or gate live cells), you can do this by setting different subets in the ggcyto command (‘root’ or ‘Singlets’).

# Gating strategy {.tabset}

## Debris gate

The axis limits are set by the instrument. Use FSC.H,y=SSC.H

```{r}
ggcyto(gs[[1]], 
       aes(x=FSC.H,y=SSC.A), subset="root") + 
  geom_hex(bins = 200) + 
  ggcyto_par_set(limits = "instrument") + 
    scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Define the gate
check the gate in a plot (before adding it to the gating set)
if the gate seems right, add / apply it to the gating set
recompute the gatingSet to have statistics available.

Its much easier to draw rectangle gates than polygon gates in R.

```{r}
# define gate
g.debris <- rectangleGate(filterId = "noDebris",
                          "FSC.H"=c(2.2e4,1e5),
                          "SSC.A"=c(2.8e4,1.3e5)) 

# check gate
ggcyto(gs[[1]],aes(x=FSC.H,y=SSC.A) ,subset="root") + 
  geom_hex(bins = 200) + 
  geom_gate(g.debris) + 
  ggcyto_par_set(limits = "instrument") + 
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gate the cells

```{r}
gs_pop_add(gs,g.debris) # add gate to GatingSet
recompute(gs)
```

```{r debris_gate, fig.height=20, fig.width=12, fig.cap = "Debris gate"}
ggcyto(gs,aes(x=FSC.H,y=SSC.A)) +
  geom_hex(bins = 200) + 
  geom_gate("noDebris") +
  geom_stats(adjust = 0.8) + 
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) + facet_wrap(~name, ncol = 3)
```

## Singlet gate

Drawing the polygon gate is a real pain, I have to manually type numbers in.

### Define the gate
check the gate in a plot (before adding it to the gating set)
if the gate seems right, add / apply it to the gating set
recompute the gatingSet to have statistics available

```{r}
# define gate
g.singlets <- polygonGate(filterId = "Singlets",
                           "FSC.A"=c(4e4,1.1e5,9.5e4,2e4),
                           "FSC.H"=c(1.5e4,9.5e4,1e5,2e4)) 

# check gate
ggcyto(gs[[1]],aes(x=FSC.A,y=FSC.H) ,subset="noDebris") + 
  geom_hex(bins = 200) + 
  geom_gate(g.singlets) + 
  ggcyto_par_set(limits = "instrument") + 
      scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gate the cells

```{r}
gs_pop_add(gs,g.singlets,
           parent="noDebris") # add gate to GatingSet

recompute(gs)
```

overview plots at each gating step.

```{r singlet_gate, fig.height=8, fig.width=12, fig.cap = "Singlet gate"}
ggcyto(gs,aes(x=FSC.A,y=FSC.H), 
       subset="noDebris") + 
  geom_hex(bins = 200) + 
  geom_gate("Singlets") +
  geom_stats(adjust = 0.8) + 
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) + facet_wrap(~name, ncol = 5)
```

## Live cell gate

DAPI is the viability dye.

Gating steps are as above. Differentiating live and dead cells solely based on DAPI exclusion.

```{r}
g.live <- rectangleGate(filterId = "Live",
                      "FSC.A"=c(3e4,1.1e5),
                      "DAPI.A"=c(2.5e2,5e3)) # define gate

# Add biexponential axis
ggcyto(gs[[4]], aes(x=FSC.A,y=DAPI.A),
       subset="Singlets") + 
  geom_hex(bins = 200) + 
  geom_gate(g.live) + 
  scale_y_flowjo_biexp() +
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

### Gating out dead cells

There must be some spectral overlap between alexa dyes and DAPI.

```{r}
gs_pop_add(gs, 
    g.live, 
    parent="Singlets") # add gate to GatingSet

recompute(gs)
```

```{r live_gate, fig.height=20, fig.width=12, fig.cap = "Live gate"}
ggcyto(gs,aes(x=FSC.A,y=DAPI.A),
       subset="Singlets") + 
  geom_hex(bins = 200) + 
  geom_gate("Live") +
  geom_stats(adjust = 0.8) + 
  scale_y_flowjo_biexp() +
  ggcyto_par_set(limits = "instrument") +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8)) +  facet_wrap(~name, ncol = 4)
```

# Set double labelled gate {.tabset}

## All samples

```{r, eval=T}
g.double <- rectangleGate(filterId="Barcode_Neg",
               "A647"=c(-1.5e2,1.5e2), 
               "A594"=c(-1.5e2,5e2))


# check gate
ggcyto(gs[[3]],aes(x=A647,y=A594) ,subset="Live") + 
  geom_hex(bins = 200) + 
  geom_gate(g.double) + 
  ggcyto_par_set(limits = "instrument") + 
  scale_x_flowjo_biexp() + scale_y_flowjo_biexp() +
  scale_colour_Publication()+ theme_Publication() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),strip.text.x = element_text(size = 8))
```

```{r fig.height=8, fig.width=12}
ggcyto(gs,
       aes(x=A647, y=A594), subset="Live") + 
  geom_hex(bins = 200) + 
  geom_gate(g.double) + 
  geom_stats(adjust = 0.8) + 
  ggcyto_par_set(limits = "instrument") + 
  scale_x_flowjo_biexp() + scale_y_flowjo_biexp() +
  theme_Publication() +
  facet_wrap(~name)
```

## Relevant samples

```{r vendor_samples, fig.height=12, fig.width=12, fig.cap = "Double labelled samples relevant vendors"}
gs_sub <- gs[c(4:12),]

ggcyto(gs_sub,
       aes(x=A647, y=A594), subset="Live") + 
  geom_hex(bins = 200) + 
  geom_gate(g.double) + 
  geom_stats(adjust = 0.8) + 
  ggtitle("") +
  ggcyto_par_set(limits = "instrument") + 
  scale_x_flowjo_biexp() + scale_y_flowjo_biexp() +
  theme_Publication() +
  facet_wrap(~sample)
```

### Cleanup
```{r}
sessionInfo()
```