---
title: "Signal to noise plots"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6,
                      fig.path='./figures/signalNoise/',
                      layout="l-body-outset")
knitr::opts_knit$set(root.dir = here::here())
library(Matrix)
library(tidyverse)
library(ggridges)
library(DropletUtils)
library(patchwork)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

## Aim

Rough analysis to visualise data and decide which plots to use for manuscript. I will make a new notebook containing a subset of useful figures to polish up for presentation in talks and manuscripts.

## Recap

Run a more challenging sample = Mouse embryonic brain E18.5
We took mouse #3, the largest one for this purpose

## Protocols

1. MULTI-Seq LMO, LMO = Capture 1
2. 10x Genomics CellPlex, CMO = Capture 2
3. Custom Nextera CMO, Nxt = Capture 3

# Read data

These files are generated in scripts/R/NN236_write_tagCounts_file.R

```{r loadData}
CMO_matrix_dir <- here::here(
  "Mouse_E18_Brain/data/Matrices", "cmo_tag_counts")

LMO_matrix_dir = here::here(
  "Mouse_E18_Brain/data/Matrices", "lmo_tag_counts")

HTO_matrix_dir = here::here(
  "Mouse_E18_Brain/data/Matrices", "nxt_tag_counts")

matrix_dir <- list(CMO_matrix_dir, LMO_matrix_dir, HTO_matrix_dir)
names(matrix_dir) <- c("CMO", "LMO", "Nxt")
```

# Read into tibble format {.tabset}

## CellPlex

```{r readCMO}
mat <- DropletUtils::read10xCounts(matrix_dir[["CMO"]])

cm <- as.data.frame(
  Matrix::as.matrix(
  t(counts(mat)
  )))
  
tb <- cm %>% 
  pivot_longer(cols = c(CMO301:CMO312),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )
tb$Protocol <- "CMO"
# Need to add 1 before log
tb$Count <- tb$Count + 1
tb_cmo <- tb
```

## MULTI-Seq LMO

```{r readLMO}
mat <-  DropletUtils::read10xCounts(matrix_dir[["LMO"]])
cm <- as.data.frame(
  Matrix::as.matrix(
  t(counts(mat)
  )))
  
tb <- cm %>% 
  pivot_longer(cols = c(MULTI_2:MULTI_13),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )
tb$Protocol <- "LMO"
# Need to add 1 before log
tb$Count <- tb$Count + 1

tb_lmo <- tb
```

## MULTI-Seq Nxt

```{r readNxt}
mat <-  DropletUtils::read10xCounts(matrix_dir[["Nxt"]])
cm <- as.data.frame(
  Matrix::as.matrix(
  t(counts(mat)
  )))
  
tb <- cm %>% 
  pivot_longer(cols = c(Nxt_451:Nxt_465),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )
tb$Protocol <- "Nxt"
# Need to add 1 before log
tb$Count <- tb$Count + 1

tb_Nxt <- tb
```

We end by saving objects

```{r save}
tb <- list(
  tb_cmo,
  tb_lmo,
  tb_Nxt
)

names(tb) <- c("CMO", "LMO" ,"Nxt")
saveRDS(tb, here::here(
  "Mouse_E18_Brain/data/Matrices", "signal_noise_counts.tb.rds"
))
```

# GGridges {.tabset}

The plot I liked most is [https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html](GGridges)

## CMO

CellPlex diluted 1:10 had the best signal to noise.

```{r plt_cmo, fig.cap="CellPlex signal to noise"}
plt_cmo <- ggplot(tb_cmo, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges() +
  scale_x_continuous(trans='log10', limits = c(1,1e5)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 18) + 
  #scale_fill_brewer(palette = "Set1") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_cmo
```

## LMO

MULTI-Seq did not work at all. Should have washed out the media properly.

```{r plt_lmo, fig.cap="MULTI-Seq signal to noise"}
plt_lmo <- ggplot(tb_lmo, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges() +
  scale_x_continuous(trans='log10', limits = c(1,1e5)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 18) + 
  #scale_fill_brewer(palette = "Dark2") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_lmo
```

## Nxt

Trimodal peak is concerning and needs further investigation. Could be cell size or doublet related.

```{r plt_Nxt, fig.cap="Custom MULTI-Seq signal to noise"}
plt_Nxt <- ggplot(tb_Nxt, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges() +
  scale_x_continuous(trans='log10', limits = c(1,1e5)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 18) + 
  #scale_fill_brewer(palette = "Dark2") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_Nxt
```

# Gather ggridges into 1 plot

The aim is to reproduce the output of the cellranger report

```{r, fig.height=30, fig.width=8, fig.cap="Summary plot"}
plt_cmo / plt_lmo / plt_Nxt
```

# Plot signal to noise for each protocol

```{r}
head_path <- "/stornext/Projects/score/GenomicsRnD/DB_NN258/multi_out/"
tail_path <- "/outs/per_sample_outs/Lox-brain-1/metrics_summary.csv"

CMO_file_path <- paste(head_path, "R010_CMO_ME", tail_path, sep="")
LMO_file_path <- paste(head_path, "R010_LMO_ME", tail_path, sep="")
Nxt_file_path <- paste(head_path, "R010_NXT_ME", tail_path, sep="")

Nxt_data = read_csv(Nxt_file_path)
Nxt_data <- Nxt_data %>% 
  filter(`Metric Name` == "CMO signal-to-noise ratio")

CMO_data = read_csv(CMO_file_path)
CMO_data <- CMO_data %>% 
  filter(`Metric Name` == "CMO signal-to-noise ratio")

LMO_data = read_csv(LMO_file_path)
LMO_data <- LMO_data %>% 
  filter(`Metric Name` == "CMO signal-to-noise ratio")

Nxt_data$Protocol <- "Custom MULTI-Seq CMO"
LMO_data$Protocol <- "MULTI-Seq LMO"
CMO_data$Protocol <- "CellPlex 1:10"

tb <- rbind(CMO_data, LMO_data, Nxt_data)
tb$`Metric Value` <- as.numeric(tb$`Metric Value`)
```

Recolour the tags to reflect the ggridges plot.
Mung the names of the tags

```{r}
tb$CellRangerCall <- tb$`Group Name`

tb$CellRangerCall <- recode(tb$CellRangerCall,
                      CMO301="Tag1",CMO302="Tag2",CMO303="Tag3",CMO304="Tag4",
                      CMO305="Tag5",CMO306="Tag6",CMO307="Tag7",CMO308="Tag8",
                      CMO309="Tag9",CMO310="Tag10",CMO311="Tag11",CMO312="Tag12",
                      Nxt_451="Tag1",Nxt_452="Tag2",Nxt_453= "Tag3",Nxt_455= "Tag4",
                      Nxt_456= "Tag5",Nxt_457= "Tag6",Nxt_458= "Tag7",Nxt_459= "Tag8",
                      Nxt_460= "Tag9",Nxt_462= "Tag10",Nxt_463= "Tag11",Nxt_465= "Tag12",
                      MULTI_2= "Tag1",MULTI_3= "Tag2",MULTI_4= "Tag3",MULTI_5= "Tag4",
                      MULTI_6= "Tag5", MULTI_7= "Tag6",MULTI_8= "Tag7",MULTI_9= "Tag8",
                      MULTI_10= "Tag9", MULTI_11= "Tag10",MULTI_12= "Tag11",MULTI_13= "Tag12"
)

tb$Protocol <- recode(tb$Protocol,
                      "Custom MULTI-Seq CMO"="Custom\nMULTI-Seq CMO")
```

## Make the plot

The signal to noise is above the recomended 5 for CellPlex diluted 1:10.  
Suprisingly based on the raw count data is the Nxt and LMO which looks different, the LMO and Nxt have similar signal to noise.

```{r signal_noise, fig.cap="Signal to noise from cell ranger"}
plt4 <- ggplot(data=tb) +
  geom_jitter(aes(y=`Metric Value`, x=Protocol, colour=CellRangerCall), 
              size=3, width = 0.15,alpha=0.66, show.legend=FALSE) +
  scale_y_continuous(breaks=seq(0,10,1)) +
  xlab("") + ylab("Signal to noise ratio") +
  scale_colour_brewer(type="qualitative", palette = "Dark2") +
  theme_Publication(base_size=16)

plt4
```


## Session info
```{r sessionInfo}
sessionInfo()
```
