---
title: "Signal to noise plots 10x Genomics example"
author: "Daniel Brown"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      smooth_scroll: true
    theme: readable
    highlight: tango 
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      dev = "png",
                      dpi = 300,
                      fig.asp=0.6,
                      fig.path='./figures/signalNoise/',
                      layout="l-body-outset")
knitr::opts_knit$set(root.dir = here::here())
library(Matrix)
library(tidyverse)
library(patchwork)
library(ggridges)
library(DropletUtils)
source(here::here("scripts/R/theme_publication.R"))
theme_set(theme_Publication())
```

# Aim

## Data description

Cells from four E18 mouse combined cortex, hippocampus and ventricular zone tissues (C57BL/6 strain) were obtained by 10x Genomics from BrainBits LLC. Cells from the four E18 mouse tissues were multiplexed at equal proportions with 3 CMOs per tissue, resulting in a pooled sample labeled with 12 CMOs. Cells from the non-multiplexed sample were used as one of the four sample types that composed the multiplexed sample.

## Libraries

Libraries were prepared following the Chromium Next GEM Single Cell 3สน Reagent Kits v3.1 (Dual Index) with Feature Barcode technology for Cell Multiplexing User Guide (CG000388).

## Preprocessing

This object is generated in notebooks/10xGenomics_example/_buildSCE.Rmd

# Read data

These files are generated in scripts/R/NN236_write_tagCounts_file.R

```{r loadData}
sce_tenX <- readRDS(here::here(
  "data/10xGenomicsExample/SCEs", "cmo-basic.rds"))

sce_DB <- readRDS(here::here(
  "data/NN265/SCEs", "cmo-anno.rds"))
```

## Read into tibble format

10x Genomics data

```{r matrixTenX, cache=TRUE}
mat <- Matrix::as.matrix(
  counts(altExp(sce_tenX
                )))

mat <- as.data.frame(t(mat))
  
tb <- mat %>% 
  pivot_longer(cols = c(CMO301:CMO312),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )

tb$Source <- "10x Genomics"
# Need to add 1 before log
tb$Count <- tb$Count + 1
tb_tenX <- tb
```

Data generated in house by myself

```{r matrixDB, cache=TRUE}
mat <- Matrix::as.matrix(
  counts(altExp(sce_DB
                )))

mat <- as.data.frame(t(mat))
  
tb <- mat %>% 
  pivot_longer(cols = c(CMO301:CMO312),
               names_to = "Multiplexing_Capture", 
               values_to = "Count"
              )

tb$Source <- "In house"
# Need to add 1 before log
tb$Count <- tb$Count + 1
tb_DB <- tb
```

# GGridges plots

The plot I liked most is [https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html](GGridges)

## 10x Genomics


```{r plt_10x, fig.cap="10x Genomics CellPlex signal to noise"}
plt_tenX <- ggplot(tb_tenX, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges() +
  scale_x_continuous(trans='log10', limits = c(1,2e5)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 18) + 
  ggtitle("10x Genomics example") +
  #scale_fill_brewer(palette = "Dark2") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_tenX
```

## In house data

```{r plt_DB, fig.cap="In house CellPlex signal to noise"}
plt_DB <- ggplot(tb_DB, aes(x = Count, y = Multiplexing_Capture, 
                   fill = Multiplexing_Capture)) +
  geom_density_ridges() +
  scale_x_continuous(trans='log10', limits = c(1,2e5)) +
  annotation_logticks(base = 10, sides = "b") +
  theme_ridges(font_size = 18) + 
  ggtitle("In house") +
  #scale_fill_brewer(palette = "Dark2") +
  ylab("") + xlab("") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(legend.position = "none")

plt_DB
```

## Combine together

```{r combo, fig.height=12, fig.width=8, fig.cap="Combined plot"}
plt_tenX / plt_DB
```

# Conclusion

The signal to noise looks similar only a shift is noticeable

## Session info
```{r sessionInfo}
sessionInfo()
```
